---
title: "Photometry data processing"
output: html_notebook
---
If you want to use this code, please cite our Jove paper:

__Martianova, E., Aronson, S., Proulx, C.D.__ [Multi-Fiber Photometry to Record Neural Activity in Freely Moving Animal.](https://www.jove.com/video/60278/multi-fiber-photometry-to-record-neural-activity-freely-moving). _J. Vis. Exp._ (152), e60278, doi:10.3791/60278 (2019).

This R notebook is a translation in R of the Python Jupyter notebook Photometry_data_processing.ipynb


# Your data
```{r}
folder = 'C:/'                             # Change to your folder
file = 'example.csv'                       # Change to your file or use our example.csv
df <- read.csv(paste(folder,file,sep=''))  # Change if the file is not *.csv 
head(df, n=5)
```

```{r}
# Change depending on your data frame
raw_reference <- df$MeanInt_410nm[2:nrow(df)]
raw_signal <- df$MeanInt_470nm[2:nrow(df)]
```


Plot raw traces
```{r}
par(mfrow=c(2,1))
plot(raw_reference, type='l', col='purple')
plot(raw_signal, type='l', col='blue')
```

# Use function get_zdFF.R to calculate z-score dF/F
```{r}
zdFF = get_zdFF(raw_reference,raw_signal)
```


Plot z-dF/F
```{r}
plot(zdFF, type='l', col='black')
```

# Analysis step-by-step

### Smooth traces


Make sure to run movavg.R function.
```{r}
smooth_win = 10
smooth_reference <- movavg(raw_reference, smooth_win)
smooth_signal <- movavg(raw_signal, smooth_win)
```
You can use other algorithms to smooth your data.


Plot smoothed traces
```{r}
par(mfrow=c(2,1))
plot(smooth_reference, type='l', col='purple')
plot(smooth_signal, type='l', col='blue')
```

### Find baseline slopes of the signals


Make sure to run airPLS.R function
```{r}
lambda = 5e4
itermax = 50
differences = 1

base_r <- airPLS(smooth_reference, lambda, differences, itermax)
base_s <- airPLS(smooth_signal, lambda, differences, itermax)
```


Plot baseline slopes
```{r}
par(mfrow=c(2,1))
plot(smooth_reference, type='l', col='purple')
lines(base_r)
plot(smooth_signal, type='l', col='blue')
lines(base_s)
```

### Remove the baselines and the begining of the traces
```{r}
remove = 200
n = length(raw_reference)
reference <- smooth_reference[remove:n] - base_r[remove:n]
signal <- smooth_signal[remove:n] - base_s[remove:n]
```


Plot traces
```{r}
par(mfrow=c(2,1))
plot(reference, type='l', col='purple')
plot(signal, type='l', col='blue')
```


### Standardize traces
```{r}
z_reference <- (reference - median(reference)) / sd(reference)
z_signal <- (signal - median(signal)) / sd(signal)
```


Plot traces
```{r}
par(mfrow=c(2,1))
plot(z_reference, type='l', col='purple')
plot(z_signal, type='l', col='blue')
```

### Fit reference trace to signal

```{r}
# Robust linear regression
require(MASS)
fit <- rlm(z_signal ~ z_reference)
```


Plot the fit
```{r}
par(mfrow=c(1,1))
plot(z_reference, z_signal)
abline(fit, col="red")
```

### Align reference trace to signal
```{r}
z_reference_fit <- predict(fit)
```


Plot aligned traces
```{r}
plot(z_signal, type='l', col='blue')
lines(z_reference_fit, type='l', col='purple')
```

### Calculate z-score zdF/F
```{r}
zdFF <- z_signal - z_reference_fit
```


Plot z-dF/F
```{r}
plot(zdFF, type='l', col='black')
```

# Contact us
If you have any questions, please contact us: ekaterina.martianova.1@ulaval.ca